\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amsfonts,amssymb,amsthm}
\usepackage{braket}

\usepackage[dvipsnames]{xcolor}
\usepackage[colorlinks, linkcolor=OliveGreen, citecolor=purple,urlcolor=purple, bookmarks=true,backref=page]{hyperref}
\usepackage[capitalise, nameinlink]{cleveref}
\usepackage{physics}
\usepackage{graphicx}
% redefine formatting of backwards references
\renewcommand{\backref}[1]{}
\renewcommand{\backreftwosep}{, }
\renewcommand{\backreflastsep}{, }
\renewcommand{\backrefalt}[4]{%
\ifcase #1 %
%
\or
[p.\ #2]%
\else
[pp.\ #2]%
\fi}

\input{new_commands.tex}
\input{math_env.tex}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Quantum Monte Carlo}

\author{Hannah Baek, Tobias Tian \\
University of British Columbia}

\date{\vspace{-2.5ex} December 8th, 2025}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\input{abstract.tex}
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Next Steps and Updated Timeline}
We highlighted in our proposal that this paper will explore the background of three variants of QMC: PIMC, DMC, and VMC. After generally reading about each of the three variants, we have decided to put our main focus on VMC, where we give explanations as detailed as possible. This is also the one on which we choose to give a code demonstration. Therefore, for the next steps, there are three main goals. The first one is to continue to refine the explanation of VMC at each critical step. The second one is to complete writing the code demonstration for VMC. The last one is to explain DMC and PIMC in a less detailed but still complete way. To summarize, what further needs to be done is:
\begin{enumerate}
    \item Refine the VMC process to explain the two remaining challenges.
    \item Complete writing the code demonstration for VMC.
    \item Explain theoretically what DMC does and how it contrasts with VMC, we will focus on \cite{troyer_11} \cite{foulkes_mitas_needs_rajagopal_01} \cite{kalos_whitlock_08}.
    \item Explore PIMC and briefly touch on the Fermion sign problem, here we will focus on \cite{troyer_11} \cite{troyer_wiese_05}.
\end{enumerate}
Our updated timeline: \\
Nov. 12 - Nov. 16: Complete writing about VMC process. Explain as detailed as possible on how trial wavefunction are chosen and optimized, and how sampling is conducted. \\
Nov. 17 - Nov. 23: Complete writing about both DMC and PIMC. Highlight on the comparison between VMC and DMC, and highlight the sign problem in PIMC. \\
Nov. 24 - Nov. 30: Complete the code demonstration for VMC. \\
Dec. 1 - Dec. 7: Finalize and refine the paper by systematically going through it checking for inconsistencies and errors.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\input{1_intro.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Monte Carlo Method}
\input{2_monte_carlo.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Variational Monte Carlo}
\input{3_variational.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diffusion Monte Carlo}
\input{4_diffusion.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Path Integral Monte Carlo}
\input{5_path_integral.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}
\input{acknowledgement.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{references}
\bibliographystyle{alphaurl}
\end{document}